### 2022-03-08 paper/output/document/appendix.tex
ugh

  Free variable:
    A = ($\lambda x. \_$)($i_1$ := False) \color{black}%
}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\begin{isamarkuptext}%
In order to make this true, we'd have to require that everything is either obligatory or
prohibited somewhere, but this makes permissibility impossible, which is clearly undesirable.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{Kroy's formalization of the FUL \label{sec: kroy_ful}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
I now implement Kroy's formalization of the Formula of Universal Law. Recall that the FUL says
``act only in accordance with that maxim which you can at the same time will a universal law'' \citep{groundwork}.
Kroy interprets this to mean that if an action is permissible for a specific agent, then it must be permissible for everyone.
This formalizes the moral intuition prohibiting free-riding. According to the categorical imperative, 
no one is a moral exception.%
\end{isamarkuptext}\isamarkuptrue%
\isacommand{abbreviation}\isamarkupfalse%
\ FUL{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}bool{\isachardoublequoteclose}\ \isakeyword{where}\ {\isachardoublequoteopen}FUL\ {\isasymequiv}\ {\isasymforall}w\ A{\isachardot}\ {\isacharparenleft}{\isacharparenleft}{\isasymexists}p{\isacharcolon}{\isacharcolon}s{\isachardot}\ {\isacharparenleft}{\isacharparenleft}P\ {\isacharbraceleft}A\ p{\isacharbraceright}{\isacharparenright}\ w{\isacharparenright}{\isacharparenright}\ \ {\isasymlongrightarrow}{\isacharparenleft}\ {\isacharparenleft}{\isasymforall}p{\isachardot}{\isacharparenleft}\ P\ {\isacharbraceleft}A\ p{\isacharbraceright}{\isacharparenright}\ w{\isacharparenright}{\isacharparenright}{\isacharparenright}\ {\isachardoublequoteclose}\isanewline
%
\isamarkupcmt{If action $A$ is permissible for some person, then, for any person $p$, action $A$ must be 
permissible for $p$. The notion of ``permissible for'' 
is captured by the substitution of $x$ for $p$.%
}%
\begin{isamarkuptext}%
This formalization does not hold in DDL, the base logic. This means that Kroy's formalization
already passes one test, and that adding it as an axiom will strengthen the logic.%
\end{isamarkuptext}\isamarkuptrue%
\isacommand{lemma}\isamarkupfalse%
\ FUL{\isacharcolon}\isanewline
\ \ \isakeyword{shows}\ FUL\isanewline
\ \ \isacommand{nitpick}\isamarkupfalse%
{\isacharbrackleft}user{\isacharunderscore}axioms{\isacharbrackright}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{oops}\isamarkupfalse%
\isanewline
%
\isamarkupcmt{\color{blue} Nitpick found a counterexample for card s = 2 and card i = 2:

@@

        The success of this lemma in Kroy's logic and the emptiness of his formalization of the FUL are 
        two errors that contribute to the failure of this test. First, the statement expressed in Step 3
        should not actually hold. Impossible actions can be permissible (do I need a citation?). For example, 
        imagine I make a trip to Target to 
        purchase a folder, and they offer blue and black folders. No one would claim that it's impermissible
        for me to purchase a red folder, or, equivalently, that I am obligated to not purchase a red folder.%
\end{isamarkuptext}\isamarkuptrue%
%
\begin{isamarkuptext}%
The second issue is that Kroy's interpretation of the formula of universal law is circular.
        His formalization interprets the FUL as prohibiting A if there is someone for whom 
        $A$'ing is not permissible. This requires some preexisting notion of the permissibility of $A$, and 
        is thus circuar. The categorical imperative is supposed to be the complete,
        self-contained rule of morality \cite{groundwork}, but Kroy's version of the FUL prescribes obligations 
        in a self-referencing manner. The FUL is supposed to define what is permissible and what isn't, 
        but Kroy defines permissibility in terms of itself.

>>>>

  Free variable:
    A = ($\lambda x. \_$)($i_1$ := False) \color{black}%
}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\begin{isamarkuptext}%
In order to make this true, we'd have to require that everything is either obligatory or
prohibited somewhere, but this makes permissibility impossible, which is clearly undesirable.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{Kroy's formalization of the FUL \label{sec: kroy_ful}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
I now implement Kroy's formalization of the Formula of Universal Law. Recall that the FUL says
``act only in accordance with that maxim which you can at the same time will a universal law'' \citep[34]{groundwork}.
Kroy interprets this to mean that if an action is permissible for a specific agent, then it must be permissible for everyone.
This formalizes the moral intuition prohibiting free-riding. According to the categorical imperative, 
no one is a moral exception.%
\end{isamarkuptext}\isamarkuptrue%
\isacommand{abbreviation}\isamarkupfalse%
\ FUL{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}bool{\isachardoublequoteclose}\ \isakeyword{where}\ {\isachardoublequoteopen}FUL\ {\isasymequiv}\ {\isasymforall}w\ A{\isachardot}\ {\isacharparenleft}{\isacharparenleft}{\isasymexists}p{\isacharcolon}{\isacharcolon}s{\isachardot}\ {\isacharparenleft}{\isacharparenleft}P\ {\isacharbraceleft}A\ p{\isacharbraceright}{\isacharparenright}\ w{\isacharparenright}{\isacharparenright}\ \ {\isasymlongrightarrow}{\isacharparenleft}\ {\isacharparenleft}{\isasymforall}p{\isachardot}{\isacharparenleft}\ P\ {\isacharbraceleft}A\ p{\isacharbraceright}{\isacharparenright}\ w{\isacharparenright}{\isacharparenright}{\isacharparenright}\ {\isachardoublequoteclose}\isanewline
%
\isamarkupcmt{If action $A$ is permissible for some person, then, for any person $p$, action $A$ must be 
permissible for $p$. The notion of ``permissible for'' 
is captured by the substitution of $x$ for $p$.%
}%
\begin{isamarkuptext}%
This formalization does not hold in DDL, the base logic. This means that Kroy's formalization
already passes one test, and that adding it as an axiom will strengthen the logic.%
\end{isamarkuptext}\isamarkuptrue%
\isacommand{lemma}\isamarkupfalse%
\ FUL{\isacharcolon}\isanewline
\ \ \isakeyword{shows}\ FUL\isanewline
\ \ \isacommand{nitpick}\isamarkupfalse%
{\isacharbrackleft}user{\isacharunderscore}axioms{\isacharbrackright}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{oops}\isamarkupfalse%
\isanewline
%
\isamarkupcmt{\color{blue} Nitpick found a counterexample for card s = 2 and card i = 2:

@@

        The success of this lemma in Kroy's logic and the emptiness of his formalization of the FUL are 
        two errors that contribute to the failure of this test. First, the statement expressed in Step 3
        should not actually hold. Impossible actions can be permissible (do I need a citation?). For example, 
        imagine I make a trip to Target to 
        purchase a folder, and they offer blue and black folders. No one would claim that it's impermissible
        for me to purchase a red folder, or, equivalently, that I am obligated to not purchase a red folder.%
\end{isamarkuptext}\isamarkuptrue%
%
\begin{isamarkuptext}%
The second issue is that Kroy's interpretation of the formula of universal law is circular.
        His formalization interprets the FUL as prohibiting A if there is someone for whom 
        $A$'ing is not permissible. This requires some preexisting notion of the permissibility of $A$, and 
        is thus circuar. The categorical imperative is supposed to be the complete,
        self-contained rule of morality, but Kroy's version of the FUL prescribes obligations 
        in a self-referencing manner. The FUL is supposed to define what is permissible and what isn't, 
        but Kroy defines permissibility in terms of itself.

### 2022-03-08 paper/output/document/thesis_2_methods.tex
ugh

Machine learning approaches like the Delphi system \citep{delphi} mentioned in Chapter \ref{intro} also may suffer explanability 
problems that my logic-programming, theorem-prover
approach does not face. Many machine learning algorithms cannot sufficiently explain their 
decisions to a human being and often find patterns in datasets that don't 
cohere with the causes that a human being would identify \citep{puiutta}. While there is significant activity 
and progress in explainable machine learning, interactive theorem provers are designed to be explainable 
at the outset. Isabelle can show the axioms and lemmas it used in constructing a proof, 
allowing a human being to reconstruct the proof independently if they wish. This is not an 
intractable problem for machine learning approaches to computational ethics, but is one reason to 
prefer logical approaches.\footnote{This argument about explanability is in the context of virtue ethics and 
machine learning. It also applies to a broader class of work in automated ethics 
that uses a ``bottom-up'' approach, in which a system learns moral judgements from prior judgements. 
I will extend this argument to general bottom-up approaches in Section \ref{relatedwork}.}%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsubsection{Kantian Ethics \label{kantianethics}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
In this paper I focus on Kantian ethics. Kant's theory is centered 
on practical reason, which is the kind of reason that we 
use to decide what to do. In \emph{The Groundwork of the Metaphysics of Morals}, Kant explains that 
rational beings are unique because we act ``in accordance with 
the representations of laws'' \citep[4:412]{groundwork}. A ball thrown into the air acts 
according to the laws of physics. It cannot ask itself, ``Should I fall back to the ground?'' 
It simply falls. A rational being, on the other hand, can ask, ``Should I act on this reason?"''
As Korsgaard describes it, when choosing which desire to act on, ``it is as if there is something over 
and above all of your desires, something which is you, and which chooses which desire to act on'' \citep[100]{sources}. 
Rational beings are set apart by this reflective capacity. We are purposive and 
our actions are guided by practical reason. We have reasons for acting, even when these reasons are
opaque to us. This reflective choosing, or operation of practical reason, is what Kant calls the will.

The will operates by adopting or willing maxims, which are its perceived reasons for acting. Kant defines a maxim as 
the ``subjective principle of willing,'' or the reason that the will \emph{subjectively} gives 
to itself for acting \citep[16 footnote 1]{groundwork}. Many philosophers agree that a maxim consists 
of some combination of circumstances, 
act, and goal.\footnote{For more discussion of the definition of a maxim, see Section \ref{whatisamaxim}.}
One example of a maxim is ``When I am hungry, I will eat a doughnut in order to satisfy my sweet tooth.''
When an agent wills this maxim, they decide to act on it. They commit themselves to the end in the maxim 
(e.g. satisfying your sweet tooth). They represent their action, to themselves, as following the 
principle given by this maxim. Because a maxim captures an agent's principle of action, Kant evaluates
maxims as obligatory, prohibited, or permissible. He argues that the form of certain maxims 
requires any rational agent to will them, and these maxims are obligatory.

@@

%:%731=567%:%
%:%732=568%:%
%:%734=569%:%
%:%735=569%:%
%:%736=570%:%
%:%737=571%:%
%:%738=571%:%
%:%739=572%:%
%:%740=573%:%
%:%741=573%:%
%:%742=574%:%
%:%744=575%:%
%:%745=576%:%
%:%748=577%:%
%:%750=578%:%
%:%751=578%:%
%:%752=579%:%
%:%754=580%:%
%:%757=585%:%
%:%758=586%:%
%:%759=587%:%
%:%760=588%:%
%:%762=591%:%
%:%763=591%:%
%:%764=591%:%
%:%766=591%:%
%:%770=591%:%
%:%771=591%:%
%:%773=592%:%
%:%774=593%:%
%:%775=593%:%
%:%777=594%:%
%:%778=595%:%
%:%779=596%:%
%:%789=598%:%
%:%790=599%:%
%:%791=600%:%

>>>>

Machine learning approaches like the Delphi system \citep{delphi} mentioned in Chapter \ref{intro} also may suffer explanability 
problems that my logic-programming, theorem-prover
approach does not face. Many machine learning algorithms cannot sufficiently explain their 
decisions to a human being and often find patterns in datasets that don't 
cohere with the causes that a human being would identify \citep{puiutta}. While there is significant activity 
and progress in explainable machine learning, interactive theorem provers are designed to be explainable 
at the outset. Isabelle can show the axioms and lemmas it used in constructing a proof, 
allowing a human being to reconstruct the proof independently if they wish. This is not an 
intractable problem for machine learning approaches to computational ethics, but is one reason to 
prefer logical approaches.\footnote{This argument about explanability is in the context of virtue ethics and 
machine learning. It also applies to a broader class of work in automated ethics 
that uses a ``bottom-up'' approach, in which a system learns moral judgements from prior judgements. 
I will extend this argument to general bottom-up approaches in Section \ref{relatedwork}.}%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsubsection{Kantian Ethics \label{kantianethics}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
Kant's theory is centered 
on practical reason, which is the kind of reason that we 
use to decide what to do. In \emph{The Groundwork of the Metaphysics of Morals}, Kant explains that 
rational beings are unique because we act ``in accordance with 
the representations of laws'' \citep[26]{groundwork}.\footnote{} A ball thrown into the air acts 
according to the laws of physics. It cannot ask itself, ``Should I fall back to the ground?'' 
It simply falls. A rational being, on the other hand, can ask, ``Should I act on this reason?"''
As Korsgaard describes it, when choosing which desire to act on, ``it is as if there is something over 
and above all of your desires, something which is you, and which chooses which desire to act on'' \citep[100]{sources}. 
Rational beings are set apart by this reflective capacity. We are purposive and 
our actions are guided by practical reason. We have reasons for acting, even when these reasons are
opaque to us. This reflective choosing, or operation of practical reason, is what Kant calls the will.

The will operates by adopting or willing maxims, which are its perceived reasons for acting. Kant defines a maxim as 
the ``subjective principle of willing,'' or the reason that the will \emph{subjectively} gives 
to itself for acting \citep[16, footnote 1]{groundwork}. Many philosophers agree that a maxim consists 
of some combination of circumstances, 
act, and goal.\footnote{For more discussion of the definition of a maxim, see Section \ref{whatisamaxim}.}
One example of a maxim is ``When I am hungry, I will eat a doughnut in order to satisfy my sweet tooth.''
When an agent wills this maxim, they decide to act on it. They commit themselves to the end in the maxim 
(e.g. satisfying your sweet tooth). They represent their action, to themselves, as following the 
principle given by this maxim. Because a maxim captures an agent's principle of action, Kant evaluates
maxims as obligatory, prohibited, or permissible. He argues that the form of certain maxims 
requires any rational agent to will them, and these maxims are obligatory.

@@

%:%731=566%:%
%:%732=567%:%
%:%734=568%:%
%:%735=568%:%
%:%736=569%:%
%:%737=570%:%
%:%738=570%:%
%:%739=571%:%
%:%740=572%:%
%:%741=572%:%
%:%742=573%:%
%:%744=574%:%
%:%745=575%:%
%:%748=576%:%
%:%750=577%:%
%:%751=577%:%
%:%752=578%:%
%:%754=579%:%
%:%757=584%:%
%:%758=585%:%
%:%759=586%:%
%:%760=587%:%
%:%762=590%:%
%:%763=590%:%
%:%764=590%:%
%:%766=590%:%
%:%770=590%:%
%:%771=590%:%
%:%773=591%:%
%:%774=592%:%
%:%775=592%:%
%:%777=593%:%
%:%778=594%:%
%:%779=595%:%
%:%789=597%:%
%:%790=598%:%
%:%791=599%:%
